Dynamic physiological behavior allows shade leaves to outperform sun leaves in high light

### 2016-05-12 gm_newphyt/manuscript_gmes.Rmd
new manuscrupt draft completed and sent to co-authors

---
title: Shade leaves of *Eucalpytus* trees lie in wait to maximize whole canopy carbon gain
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  pdf_document: default
  html_document:
    number_sections: yes
csl: new-phytologist.csl
bibliography: references.bib
---

@@

\
Author for Correspondence:
\
*Courtney Campany*
\
*Tel: +61(0)432391114*
\
*Email: courtneycampany@gmail.com*

\
```{r wc, echo=FALSE, warning=FALSE}
library(pixiedust)
wc_tab <- data.frame(count = c("Total word count (main body):", "Summary:", "Introduction:", "Materials and Methods:", "Results:",                               "Discussion:", "Acknowledgements:"), 
                     answer = c(5535, 219, 1207, 1847, 1102, 1359,20),
                     count2 = c("No. of Figures:", "No of Tables:", "No. of Supporting Information files:"
                                ,"","","",""),
                     answer2 = c("6 (all in color)",2,"6 (5 Figures, 1 Methods)", "","","",""))
#colnames(wc_tab) <- NULL 
dust(wc_tab)%>%
sprinkle_print_method("markdown")
#Dynamic respones of shade leaf physiology to light availability increases whole tree carbon gain
```

@@

# Summary
* Light gradients within tree canopies play a major role in the distribution of plant resources that define leaf photosynthetic capacity. A lack of empirical data relating photosynthesis to leaf physiological behavior within tree canopies, however, impedes our ability to assess the contribution of shade leaves to canopy carbon gain.
* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
* Compared to sun leaves, shade leaves had lower A, g~m~, leaf nitrogen and photosynthetic biochemical parameters (Vc~max~ and J~max~), but g~s~ was similar. When light intensity was increased for shade leaves both g~s~ and g~m~ increased, leading to increases in A greater than sun leaves. 
* Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain. We argue that the up-regulation of g~m~ over short time intervals with light enables shade leaves to respond quickly to sunflecks, possibly representing a new mechanism underpinning leaf gas exchange responses to light.

@@

#Introduction
Light availability is one of the most important environmental drivers of leaf carbon (C) uptake in trees. Predicting C uptake of forests usually involves upscaling leaf level measurements to assess whole canopy function. Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple upscaling based on solar irradiance problematic. Incident photosynthetic photon flux density (PPFD) declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor]. Consequently, structural and functional properties of leaves are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance strongly affects rates of leaf photosynthesis (*A~n~*) [@evans1995carbon]. To estimate whole canopy C gain it is thus necessary to account for the non-linear response of *A~n~* to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].

@@

\
Climate warming may also affect the physiological behaviour of leaves within a canopy. This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) vary with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis]. How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for upscaling *A~n~* for sun and shade leaves. Additionally, light saturated rates of *A~n~* are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature]. How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients. The impacts of warming on plant physiological processes are vast, yet differentiating their impacts on leaf physiology within a canopy will be essential to evaluate whole tree responses to a changing climate.

@@

##Conclusions
Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements to the canopy. Although resource allocation constrains leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain. These findings suggest that current theories of leaf optimal behaviour should be extended to include dynamic light environments, which will have important implications for process-based models that predict canopy C gain from rates of leaf photosynthesis. Furthermore, the dynamic nature of g~m~ cannot be simply parameterized in tree growth models and possibly should be excluded until it can be represented properly. Additional empirical data, across multiple tree species, are needed to determine both the mechanisms and the capacity of g~m~ to opportunistically increase CO~2~ drawdown. To improve our ability to predict whole canopy C gain, future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.

@@

**Table 1**. *Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments. Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6). Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light. Units of LMA and Leaf N~area~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, $\Psi$ is MPa and $\delta$^13^C is &permil;. Different letters represent significant differences between leaf type and temperature treatments. The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.

@@

```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/par2.R")
```  
\
**Figure 1**. Bars represent the local light environment for sun and shade leaves during six gas exchange campaigns from October 2013 to April 2014. Means &pm; 1 standard error represent integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf. Each date represents the starting date for each measurement campaign. Points represent the mean (&pm; 1 standard error) daily maximum air temperature during each campaign period.

\
```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
 source("master_scripts/aci_vcmaxnitro.R")
```

\
**Figure 2**. (a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments. ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^). (b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light. (c) The relationship between *A~n~* and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions. For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).
\
```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/Agmgs.R")
```  
\
**Figure 3**. The relationship between *A~n~* to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments. Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b). Grey areas are 95% confidence intervals for the mean.
\
```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/physiology_boxplots.R")
```  
\
**Figure 4**. Box plots of measured g~s~ (a), g~m~ (b) and *A~n~* (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
\
```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=600}
source("master_scripts/wateruse.R")
```  
\
**Figure 5**. (a) The relationship between instantaneous transpiration efficiency (ITE) and VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments. (b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves at high light and shade leaves at low light. For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment. For (b) the dashed line represents the significant linear model fit across all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.
\
```{r gmshade, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/gmes_shadeleaves_figure.R")
```  
\
**Figure 6**. The change in measured g~m~ for individual shade leaves following an increase in PPFD to match the light environment of full sun leaves in the same tree. Measurements of g~m~ were recorded once CO~2~ and water vapour fluxes were stable in the leaf cuvette, which took approximately 25 minutes after light intensity was increased. The dashed line is the 1:1 relationship.
\
```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/cicc_boxplots.R")
```  
\
**Figure 7**. Boxplots of (a) intercellular CO~2~ concentration (C~i~), (b) CO~2~ concentration in the chloroplasts (C~c~) and (c) CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light (C~i~-C~c~).

@@

Here, g~m~ is expressed in units of mol m^-2^ s^-1^ bar^-1^ to be consistent with Evans & Von Caemmerer [-@evans2013temperature]. As the pressure term is nearly equal to unity, suitable comparisons between g~m~ and g~s~ can be made.

>>>>

---
title: Dynamic physiological behavior allows shade leaves to outperform sun leaves in high light
author: ""
date: ""
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  pdf_document: default
  html_document:
    number_sections: yes
csl: new-phytologist.csl
bibliography: references.bib
---

@@

\
Author for Correspondence:
\
*Courtney Campany*
\
*Tel: +61(0)432391114*
\
*Email: courtneycampany@gmail.com*
\
\
Total word count (excluding summary, references and legends): 5563
\
Summary: 200
\
Introduction: 1186
\
Materials and Methods: 1863
\
Results: 1098
\
Discussion: 1380
\
Acknowledgements: 36
\
Number of figures: 7 (6 in color)
\
Number of tables: 2
\
Number of supporting information files: 4 figures, 1 methods

@@

# Summary
* Light gradients within tree canopies play a major role in the distribution of plant resources that define leaf photosynthetic capacity. A lack of empirical data relating photosynthesis to leaf physiological behavior within tree canopies, however, impedes our ability to assess the contribution of shade leaves to canopy carbon gain.
* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure net leaf photosynthesis (*A~n~*), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
* Compared to sun leaves, shade leaves had lower A, g~m~, leaf nitrogen and photosynthetic biochemical parameters (Vc~max~ and J~max~), but g~s~ was similar. When light intensity was increased for shade leaves both g~s~ and g~m~ increased, leading to increases in A greater than sun leaves. 
* Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for up-scaling leaf level measurements and predicting whole canopy carbon gain. We argue that the up-regulation of g~m~ over short time intervals with light enables shade leaves to respond quickly to sunflecks, possibly representing a new mechanism underpinning leaf gas exchange responses to light.

@@

#Introduction
Light availability is one of the most important environmental drivers of leaf carbon (C) uptake in trees. Predicting C uptake of forests usually involves up-scaling leaf level measurements to assess whole canopy function. Due to the costs and limitations of efficient light harvesting, plants cannot expose all leaves to full sun [@niinemets2010review], making simple up-scaling based on solar irradiance problematic. Incident photosynthetic photon flux density (PPFD) declines exponentially with cumulative leaf area index, creating a steep light gradient from the canopy top to bottom [@monsi2005factor]. Consequently, structural and functional properties of leaves are modified to efficiently use light [@vogelman1996leaves; @valladares2004photosynthetic], as changing irradiance strongly affects rates of leaf photosynthesis (*A~n~*) [@evans1995carbon]. To estimate whole canopy C gain it is thus necessary to account for the non-linear response of *A~n~* to light by distinguishing between shaded and sunlit leaves [@de1997simple; @linderson2012up].

@@

\
Climate warming may also affect the physiological behaviour of leaves within a canopy. This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) vary with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis]. How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for up-scaling *A~n~* for sun and shade leaves. Additionally, light saturated rates of *A~n~* are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature]. How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients. The impacts of warming on plant physiological processes are vast, yet differentiating their impacts on leaf physiology within a canopy will be essential to evaluate whole tree responses to a changing climate.

@@

##Conclusions
Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for up-scaling leaf level measurements to the canopy. Although resource allocation constrains leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain. These findings suggest that current theories of leaf optimal behaviour should be extended to include dynamic light environments, which will have important implications for process-based models that predict canopy C gain from rates of leaf photosynthesis. Furthermore, the dynamic nature of g~m~ cannot be simply parameterized in tree growth models and possibly should be excluded until it can be represented properly. Additional empirical data, across multiple tree species, are needed to determine both the mechanisms and the capacity of g~m~ to opportunistically increase CO~2~ drawdown. To improve our ability to predict whole canopy C gain, future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.

@@

**Table 1**. *Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments. Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K~l~ values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6). Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light. Units of LMA and leaf N~a~ are g m^2^, K~l~ is mmol m^-2^ s^-1^ MPa, $\Psi$ is MPa and $\delta$^13^C is &permil;. Different letters represent significant differences between leaf type and temperature treatments. The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.

@@

```{r par, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,dev="png",message=FALSE, dpi=600, dev.args=list(type="cairo")}
source("master_scripts/par2.R")
```  
\
**Figure 1**. Bars represent the local light environment for sun and shade leaves during six gas exchange campaigns from October 2013 to April 2014. Means &pm; 1 standard error represent integrated PPFD, measured with a ceptometer, at the canopy height of each selected leaf. Each date represents the starting date for each measurement campaign. Points represent the mean (&pm; 1 standard error) daily maximum air temperature during each campaign period.

\
```{r acicurves, fig.width=8, fig.height=10, warning=FALSE,echo=FALSE,dev="png", dpi=600,message=FALSE, dev.args=list(type="cairo")}
 source("master_scripts/aci_vcmaxnitro.R")
```

\
**Figure 2**. (a) ACi curves for sun and shade leaves at elevated (ET) and ambient (AT) temperature treatments. ACi curves were measured once on all trees, in February 2014, at 25°C and at saturating light (1800 $\mu$mols m^-1^ s^-1^). (b) The relationship between Vc~max~ and mean leaf N~a~ for each chamber, including sun leaves and shade leaves at low light. (c) The relationship between *A~n~* and leaf N~a~ for sun and shade leaves measured under their ambient light and temperature conditions. For (b,c) the dashed line represents the significant linear model fit for all leaves, with a marginal and conditional R^2^ of 0.28 and 0.35 for (b) and 0.24 and 0.33 for (c).
\
```{r Agmgs, fig.width=7, fig.height=7, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/Agmgs.R")
```  
\
**Figure 3**. The relationship between *A~n~* to g~s~ (a) and g~m~ (b) for sun leaves measured at high light and shade leaves measured at both low and high light under their respective elevated and ambient temperature treatments. Lines represent either smoothed regressions from a generalized additive model fit (a) or linear model fits (b). Grey areas are 95% confidence intervals for the mean.
\
```{r barplots, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600,message=FALSE, dev.args=list(type="cairo")}
source("master_scripts/physiology_boxplots.R")
```  
\
**Figure 4**. Box plots of measured g~s~ (a), g~m~ (b) and *A~n~* (c) of sun leaves and shade leaves at both low and high light pooled across six measurement dates.
\
```{r wateruse, fig.width=7, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE,dev="png", dpi=600}
source("master_scripts/wateruse.R")
```  
\
**Figure 5**. (a) The relationship between instantaneous transpiration efficiency (ITE) and VPD for sun leaves and shade leaves at both low and high light with elevated and ambient temperature treatments. (b) The relationship between leaf $\delta$^13^C and leaf N~a~ for sun leaves at high light and shade leaves at low light. For (a) VPD is the leaf to air pressure difference inside the gas exchange cuvette and lines represent predictions from the optimal ITE model with a g~1~ value for each leaf type and treatment. For (b) the dashed line represents the significant linear model fit across all leaves with a marginal and conditional R^2^ of 0.41 and 0.45, respectively.
\
```{r gmshade, fig.width=6, fig.height=6, echo=FALSE,warning=FALSE,dev="png", dpi=600,message=FALSE, dev.args=list(type="cairo")}
source("master_scripts/gmes_shadeleaves_figure.R")
```  
\
**Figure 6**. The change in measured g~m~ for individual shade leaves following an increase in PPFD to match the light environment of full sun leaves in the same tree. Measurements of g~m~ were recorded once CO~2~ and water vapour fluxes were stable in the leaf cuvette, which took approximately 25 minutes after light intensity was increased. The dashed line is the 1:1 relationship.
\
```{r cicc, fig.width=10, fig.height=6, echo=FALSE,warning=FALSE,message=FALSE,dev="png", dpi=600, dev.args=list(type="cairo")}
source("master_scripts/cicc_boxplots.R")
```  
\
**Figure 7**. Boxplots of (a) intercellular CO~2~ concentration (C~i~), (b) CO~2~ concentration in the chloroplasts (C~c~) and (c) CO~2~ drawdown from substomatal cavities to sites of carboxylation of sun leaves and shade leaves at both low and high light (C~i~-C~c~).

@@

Here, g~m~ is expressed in units of mol m^-2^ s^-1^ bar^-1^ to be consistent with Evans & von Caemmerer [-@evans2013temperature]. As the pressure term is nearly equal to unity, suitable comparisons between g~m~ and g~s~ can be made.
